Practice Problems

### 2023-09-28 Stats Rethinking/Ch8/Practice Problems.qmd
Finished through medium

1.  Bread dough rises because of yeast.
    1.  Temperature - Too cold or hot an the bacteria die, somewhere there is a sweet-spot for it
2.  Education leads to higher income.
    1.  
3.  Gasoline makes a car go.

@@

1.  Caramelizing onions requires cooking over low heat and making sure the onions do not dry out.
2.  A car will go faster when it has more cylinders or when it has a better fuel injector.
3.  Most people acquire their political beliefs from their parents, unless they get them instead from their friends.
4.  Intelligent animal species tend to either be highly social or have manipulative appendages (hands, tentacles, etc.)

# 8E3

@@

@@

@@

@@

@@

>>>>

1.  Bread dough rises because of yeast.
    1.  Temperature - Too cold or hot an the bacteria die, somewhere there is a sweet-spot for it
2.  Education leads to higher income.
    1.  Gender
3.  Gasoline makes a car go.
    1.  Horsepower

@@

1.  Caramelizing onions requires cooking over low heat and making sure the onions do not dry out.
    1.  No interaction
2.  A car will go faster when it has more cylinders or when it has a better fuel injector.
    1.  Interaction
3.  Most people acquire their political beliefs from their parents, unless they get them instead from their friends.
    1.  No interaction
4.  Intelligent animal species tend to either be highly social or have manipulative appendages (hands, tentacles, etc.)
    1.  Interaction

# 8E3\*

@@

1\)

$$
O_i\sim Normal(\mu_i, \sigma)\\
\mu_i=\alpha+\beta(H_i-\bar{H})\\
\alpha\sim Normal(1,1)\\
\beta\sim Normal(1,1)\\
\sigma\sim Exponential(1)
$$

2\)

$$
S_i\sim Normal(\mu_i, \sigma)\\
\mu_i=\alpha+(\gamma_{C,i})C_i+\beta_{FI}FI_i\\
\gamma_{C,i}=\beta_C+\beta_{CFI}FI_i\\
\alpha\sim Normal(0.5,1)\\
\beta_C\sim Normal(0,1)\\
\beta_{FI}\sim Normal(0,1)\\
\sigma\sim Exponential(1)
$$

3\)

I have no idea how to model this

$$
P_i\sim Normal(\mu_i,\sigma)\\
\mu_i=\alpha+\beta
$$

4\)

I have no idea how to model this one either

@@

Could we just change the equation to $\mu_i=C_i(\alpha+\gamma_{W,i}W_i+\beta_SS_i)$ where $\gamma_{W,i}=\beta_W+\beta_{WS}S_i$ and $C_i$ is a dummy which represents if it was cold in the room (1 yes, 0 no). This would make the model usually predict a 0 for the buds but then could leave a possibility of it blooming.

@@

See 8M1.

@@

```{r}
data("tulips")
d <- tulips

d$blooms_std <- d$blooms / max(d$blooms)
d$water_cent <- d$water - mean(d$water)
d$shade_cent <- d$shade - mean(d$shade)

a <- rnorm(1e4, 0.5, 1); sum(a<0 | a>1)/length(a)
a <- rnorm(1e4, 0.5, 0.25); sum(a<0 | a>1)/length(a)
```

```{r}
m8.5 <- quap(
  alist(
    blooms_std ~ dnorm(mu, sigma),
    mu <- bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dnorm(0, 0.25),
    bs ~ dnorm(0, 0.25),
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ), data=d
)

m8.5.2 <- quap(
  alist(
    blooms_std ~ dnorm(mu, sigma),
    mu <- bw*water_cent + bs*shade_cent + bws*water_cent*shade_cent,
    a ~ dnorm(0.5, 0.25),
    bw ~ dnorm(0.5, 0.25),
    bs ~ dnorm(-0.5, 0.25),
    bws ~ dnorm(0, 0.25),
    sigma ~ dexp(1)
  ), data=d
)
```

```{r}
precis(m8.5)
```

```{r}
precis(m8.5.2)
```

```{r}
compare(m8.5, m8.5.2, func=WAIC)
```

```{r}
compare(m8.5, m8.5.2, func=PSIS)
```

Doesn't appear that they change the interaction effect at all.

@@

Return to the `data(tulips)` example in the chapter. Now include the `bed` variable as a predictor in the interaction model. Don't interact `bed` with the other predictors; just include it a s a main effect. Not that `bed` is categorical. So to use it properly, you will need to either construct dummy variables or rather an index variable, as explained in Chapter 5.
